---
title: "draft"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#grades in 1994-1995 school year 
School_perf1 <- allwave1%>%
  select(AID, H1ED11, H1ED12, H1ED13, H1ED14)%>%
  filter( H1ED11!= 99 ,H1ED11!= 5, H1ED11!= 6)%>%
  filter( H1ED12!= 99 ,H1ED12!= 5, H1ED12!= 6)%>%
  filter( H1ED13!= 99 ,H1ED13!= 5, H1ED13!= 6)%>%
  filter( H1ED14!= 99 ,H1ED14!= 5, H1ED14!= 6)%>%
    na_if(96)%>%na_if(97)%>%na_if(98)

#pattern of missing data
md.pattern(School_perf1)

#number of missing data
sapply(School_perf1, function(x) sum(is.na(x))) # 10 percent of the data is missing

#transform into factor
School_perf1 <- School_perf1 %>%
  mutate( H1ED11 = as.factor(H1ED11),
          H1ED12 = as.factor(H1ED12),
          H1ED13 = as.factor(H1ED13),
          H1ED14 = as.factor(H1ED14)
          )
#impute missing data
tempData <- mice(School_perf1, m =5, maxit = 5, method = "polr", seed = 500)
summary(tempData)

#completed data
completed_data <- complete(tempData,1)
md.pattern(completed_data)

```


```{r}

#grades in 1996 school  year
School_perf2 <- allwave2%>%
  select(AID, H2ED7, H2ED8, H2ED9, H2ED10)%>%
  filter(H2ED7!= 5, H2ED7!= 6)%>%
  filter( H2ED8 != 5, H2ED8!= 6)%>%
  filter( H2ED9 != 5, H2ED9 != 6)%>%
  filter( H2ED10 != 5, H2ED10 != 6)%>%
    na_if(96)%>%na_if(97)%>%na_if(98)

#pattern of missing data
md.pattern(School_perf2)

#number of missing data
sapply(School_perf2, function(x) sum(is.na(x)))

#transform into factor
School_perf2 <- School_perf2 %>%
  mutate( H2ED7 = as.factor(H2ED7),
          H2ED8 = as.factor(H2ED8),
          H2ED9 = as.factor(H2ED9),
          H2ED10 = as.factor(H2ED10)
          )
#impute missing data
tempData2 <- mice(School_perf2, m =5, maxit = 5, method = "polr", seed = 500)
summary(tempData2)

#completed data
completed_data2 <- complete(tempData2,1)
md.pattern(completed_data2)

nrow(School_perf1) #15942
nrow(School_perf2) # 10663


```

```{r}
#Drop in performance
#GET AID of individual present in both wave 1 and wave 2
merged<-left_join(completed_data, completed_data2, by = "AID")%>%
   mutate( H2ED7 = as.numeric(H2ED7),
          H2ED8 = as.numeric(H2ED8),
          H2ED9 = as.numeric(H2ED9),
          H2ED10 = as.numeric(H2ED10),
          H1ED11 = as.numeric(H1ED11),
          H1ED12 = as.numeric(H1ED12),
          H1ED13 = as.numeric(H1ED13),
          H1ED14 = as.numeric(H1ED14))%>%
   mutate(ENG = H1ED11 - H2ED7 +1, MTH = H1ED12 - H2ED8 +1, HST = H1ED13 - H2ED9 +1, SCI = H1ED14 - H2ED10+1)%>%
  mutate(PERF = ENG + MTH +HST + SCI)
  
  
nrow(merged) # (15942)
table(merged$PERF)



perf_drop <- merged %>%
  select (AID, PERF)%>%
  mutate(PERF = as.factor(ifelse(perf_drop$PERF<=0,1,0 )))
         
table(perf_drop$PERF)
md.pattern(perf_drop)

wave1_ID <- allwave1 %>%
  select(AID)

wave1drop_performance <- left_join(wave1_ID, perf_drop, by ="AID")
md.pattern(wave1drop_performance)


```






```{r}


#scaled answers
learnt_about_suicide <- allwave1%>%
  select(AID,H1TS17)%>%
  na_if(6)%>%na_if(8)

#turn into factors

learnt_about_suicide  <- learnt_about_suicide  %>%
    mutate(H1TS17 = as.factor(H1TS17))

#pattern of missing data
md.pattern(learnt_about_suicide)

#Cannot impute data no predictors


```





```{r}
#scaled attempt
suicide_sc <- allwave1%>%
  select(AID, H1SU2)%>%
  na_if(6)%>%na_if(7)%>%na_if(8)%>%
  drop_na()

#2743 obs on attempt on suicide
```







```{r}

#Negativity_feeling
neg_feeling <- allwave1%>%
  select(AID, H1FS3, H1FS6, H1FS9, H1FS13,H1FS16, H1FS19)%>%
  na_if(6)%>%na_if(8)

#missing data 
md.pattern(neg_feeling)  # 357 missing data in total

#turning into factor

neg_feeling <- neg_feeling %>%
  mutate(H1FS3 = as.factor(H1FS3),
         H1FS6 = as.factor(H1FS6),
         H1FS9 = as.factor(H1FS9),
         H1FS13 = as.factor(H1FS13),
         H1FS16 = as.factor(H1FS16),
         H1FS19 = as.factor(H1FS19))
#imputing missing data 

tempData_nf <- mice(neg_feeling, m =5, maxit = 5, method = "polr", seed = 500)
completed_nf <- complete(tempData_nf,1)

#checking
md.pattern(completed_nf)
```

```{r}
#positive_feeling

pos_feeling <- allwave1%>%
  select(AID, H1FS11, H1FS15)%>%
  na_if(6)%>%na_if(8)

#missing data
md.pattern(pos_feeling)

#turn into factor
pos_feeling <- pos_feeling%>%
  mutate(H1FS11 = as.factor(H1FS11), 
         H1FS15 = as.factor(H1FS15))
#imputing missing data 
tempData_pf <- mice(pos_feeling, m =5, maxit = 5, method = "polr", seed = 500)
completed_pf <- complete(tempData_pf,1)

#Checking 
md.pattern(completed_pf)

```
```{r}
logModel0 <- glm(H1SU1~ H1ED16 + H1ED17 + H1ED18 + H1ED19 + H1ED20, family = binomial(link = "logit"), data = training)
print(summary(logModel0))

anova(logModel0, test = "Chisq") #Sex , H1ED7

print(summary(logModel))
logModel <- glm(H1SU1~ BIO.SEX + AGE + H1ED5 + H1ED7 + H1ED9, family = binomial(link = "logit"), data = training)

print(summary(logModel))

anova(logModel, test = "Chisq") #Sex , H1ED7

logModel1 <- glm(H1SU1~ H1GH2 + H1GH4 + H1GH6 + H1GH8 + H1GH11 + H1GH18 , family = binomial(link = "logit"), data = training)

print(summary(logModel1))

anova(logModel1, test = "Chisq") # 2, 6, 18

logModel2 <- glm(H1SU1~ BIO.SEX + H1ED7 + H1GH2   + H1GH6 + H1GH18 , family = binomial(link = "logit"), data = training)

anova(logModel2, test = "Chisq")

logModel3 <- glm(H1SU1~ BIO.SEX  + H1GH2 + H1GH6 + H1GH18 + H1ED16, family = binomial(link = "logit"), data = training)
print(summary(logModel3))
anova(logModel3, test = "Chisq")

logModel4 <- glm(H1SU1~    shake_blues + depressed + failure + Happy + lonely + enjoy_life + sad + not_worth_living , family = binomial(link = "logit"), data = training)
print(summary(logModel4))
anova(logModel4, test = "Chisq") #shake_blues, depressed, failures, not worth living

logModel41 <- glm(H1SU1~    shake_blues + depressed + failure + not_worth_living , family = binomial(link = "logit"), data = training)
print(summary(logModel41))
anova(logModel41, test = "Chisq") #shake_blues, depressed, failures, not worth living


logModel5 <- glm(H1SU1~    alcohol + cigarette + marijuana + cocaine + inhalant + illegal_drug + PERF, family = binomial(link = "logit"), data = training)
print(summary(logModel5))
anova(logModel5, test = "Chisq") #alcohol, cigarette, marijuana, inhalant

logModel6 <- glm(H1SU1~ BIO.SEX  + H1GH2 + H1GH6 + H1GH18 + H1ED16 +alcohol + cigarette + marijuana + inhalant +shake_blues + depressed + failure + not_worth_living, family = binomial(link = "logit"), data = training)
print(summary(logModel6))
anova(logModel6, test = "Chisq")

logModel7 <- glm(H1SU1~ BIO.SEX  + H1GH2 + H1GH6 + H1GH18 + H1ED16 +alcohol + inhalant +shake_blues + depressed + failure + not_worth_living, family = binomial(link = "logit"), data = training)
print(summary(logModel7))
anova(logModel7, test = "Chisq")

logModel8 <- glm(H1SU1~ BIO.SEX  + H1GH2 + H1GH6 + H1GH18 + H1ED16 +alcohol + inhalant +shake_blues + depressed + failure + not_worth_living + H1SU4.y , family = binomial(link = "logit"), data = training)
print(summary(logModel8))
anova(logModel8, test = "Chisq")

logModel9 <- glm(suicidal_thoughts~ . , family = binomial(link = "logit"), data = training)
print(summary(logModel9))
anova(logModel8, test = "Chisq")


```
#asessing predictive ability
```{r}
fitted.results <- predict(logModel2, newdata = testing, type = 'response')
fitted.results <- as.matrix(fitted.results)
fitted.results <- ifelse(fitted.results>0.5, 1,0)
error <- mean(fitted.results == testing$H1SU1)
print(paste('Logistic Regression Accuracy', error))

fitted.result1 <- predict(logModel3, newdata = testing, type = 'response')
fitted.result1 <- as.matrix(fitted.result1)
fitted.result1 <- ifelse(fitted.result1>0.5, 1,0)
error1 <- mean(fitted.result1 == testing$H1SU1)
print(paste('Logistic Regression Accuracy', error1))

fitted.result2 <- predict(logModel7, newdata = testing, type = 'response')
fitted.result2 <- as.matrix(fitted.result2)
fitted.result2 <- ifelse(fitted.result2>0.5, 1,0)
error2 <- mean(fitted.result2 == testing$H1SU1)
print(paste('Logistic Regression Accuracy', error2))

fitted.result3 <- predict(logModel8, newdata = testing, type = 'response')
fitted.result3 <- as.matrix(fitted.result3)
fitted.result3 <- ifelse(fitted.result3>0.5, 1,0)
error3 <- mean(fitted.result3 == testing$H1SU1, na.rm = TRUE)
print(paste('Logistic Regression Accuracy', error3))
```

```{r}
logModel0 <- glm(suicidal_thoughts~ headache + stomach_ache + physically_weak + tired + wake_up_tired + poor_appetite + trouble_sleeping + moodiness + risk_seatbelt + risk_drunk_drive + miss_school + night_away, family = binomial(link = "logit"), data = training)
print(summary(logModel0))
an0<-anova(logModel0, test = "Chisq")
sum(an0$Deviance, na.rm = TRUE)

logModel1 <- glm(suicidal_thoughts~ repeat_grade + suspension + expelled + pay_attention + homework_done + get_along_with_students + feel_close_to_people + feel_part_of_your_school, family = binomial(link = "logit"), data = training)
print(summary(logModel1))
an1<-anova(logModel1, test = "Chisq")
sum(an1$Deviance, na.rm = TRUE)

logModel2 <- glm(suicidal_thoughts~ suicidal_surrounding + BIO.SEX + AGE + alcohol + cigarette + marijuana + cocaine + inhalant + illegal_drug , family = binomial(link = "logit"), data = training)
print(summary(logModel2))
an2 <- anova(logModel2, test = "Chisq")
sum(an2$Deviance, na.rm = TRUE)

logModel3 <- glm(suicidal_thoughts~ poor_appetite_feel + shake_blues + depressed + failure + Happy + lonely + enjoy_life + sad + not_worth_living , family = binomial(link = "logit"), data = training)
print(summary(logModel3))
an3<- anova(logModel3, test = "Chisq")
sum(an3$Deviance, na.rm = TRUE)


logModel4 <- glm(suicidal_thoughts~ suicidal_surrounding + poor_appetite_feel + shake_blues + depressed + failure+ not_worth_living + suicidal_surrounding + BIO.SEX + alcohol + cigarette + marijuana + pay_attention + feel_close_to_people +  headache + stomach_ache + physically_weak +poor_appetite + trouble_sleeping + moodiness+enjoy_life + inhalant , family = binomial(link = "logit"), data = training)
print(summary(logModel4))
an4 <- anova(logModel4, test = "Chisq")
sum(an4$Deviance, na.rm = TRUE)

logModel5 <- glm(suicidal_thoughts~ suicidal_surrounding + poor_appetite_feel + shake_blues + depressed + failure+ not_worth_living + suicidal_surrounding + alcohol , family = binomial(link = "logit"), data = training)
print(summary(logModel5))
an5 <- anova(logModel5, test = "Chisq")
sum(an5$Deviance, na.rm = TRUE)

logModel8 <- glm(suicidal_thoughts~ BIO.SEX  + headache + physically_weak + trouble_sleeping + pay_attention +alcohol + inhalant +shake_blues + depressed + failure + not_worth_living + suicidal_surrounding + poor_appetite_feel , family = binomial(link = "logit"), data = training)
print(summary(logModel8))
an<-anova(logModel8, test = "Chisq")

sum(an$Deviance, na.rm = TRUE)


```

```{r}
fitted.results <- predict(logModel5, newdata = testing, type = 'response')
fitted.results <- as.matrix(fitted.results)
fitted.results <- ifelse(fitted.results>0.5, 1,0)
error <- mean(fitted.results == testing$suicidal_thoughts, na.rm = TRUE)
print(paste('Logistic Regression Accuracy', error))

fitted.results1 <- predict(logModel8, newdata = testing, type = 'response')
fitted.results1 <- as.matrix(fitted.results1)
fitted.results1 <- ifelse(fitted.results1>0.5, 1,0)
error1 <- mean(fitted.results1 == testing$suicidal_thoughts, na.rm = TRUE)
print(paste('Logistic Regression Accuracy', error1))

fitted.results1 <- predict(logModel4, newdata = testing, type = 'response')
fitted.results1 <- as.matrix(fitted.results1)
fitted.results1 <- ifelse(fitted.results1>0.5, 1,0)
error1 <- mean(fitted.results1 == testing$suicidal_thoughts, na.rm = TRUE)
print(paste('Logistic Regression Accuracy', error1))

print("Confusion Matrix for Logistic regression");table(testing$suicidal_thoughts, fitted.results1>0.5)


True_negative2 <- 5203
True_Positive2<-  178
False_negative2<-  641
False_positive2<-  105
Actual_true2 <- 819
Actual_false2 <- 5308
predicted_true2 <- 283
Accuracy2 <- round((True_Positive2+True_negative2)/6149, digits = 5)
print(paste('Overall, the predictions are correct', Accuracy2*100,"% of time" ))
error_rate2 <- round((False_negative2 + False_positive2)/6149, digits =2)
print(paste('Overall, the predictions are wrong', error_rate2*100,"% of time" ))
True_positive_rate2 <- round(True_Positive2/Actual_true2, digits = 2)
print(paste('When students actually had suicidal thoughts, it predicts them correctly', True_positive_rate2*100,"% of time" ))
False_positive_rate2 <- round(False_positive2/Actual_false2, digits = 2)
print(paste('When students did not actually have suicidal thoughts, it predicts  they did', False_positive_rate2*100,"% of time" ))
True_negative_rate2<- round(True_negative2/Actual_false2, digits = 2)
print(paste('When students did not actually have suicidal thoughts, it predicts they did not', True_negative_rate2*100,"% of time" ))
precision2 <- round(True_Positive2/predicted_true2,digits =2)
print(paste('When the model predicts suicidal thoughts, it is correct', precision2*100,"% of time" ))
                          


```

#Confusion Matrix

```{r}

print("Confusion Matrix for Logistic regression");table(testing$H1SU1, fitted.result2>0.5)

                          

```
There were 6149 cases. 23 were predicted to have had suicidal thoughts, 6136 were predicted they don't.
In reality 824 actually had suicidal thoughts and 5325 did not have suicidal thoughts
#0
True negative: 5316
True Positive: 4
False negative: 820
False positive: 9
#2
True_negative2 <- 5211
True_Positive2<-  196
False_negative2<-  628
False_positive2<-  114
Actual_true2 <- 824
Actual_false2 <- 5325
predicted_true2 <- 310
Accuracy:

#3
#tree
```{r}
tree <- ctree(H1SU1~ BIO.SEX  + H1GH2 + H1GH6 + H1GH18 + H1ED16, training)
plot(tree, type = 'simple')

tree1 <- ctree(H1SU1~ BIO.SEX  + H1GH2 + H1GH6 + H1GH18 + H1ED16 +alcohol + inhalant +shake_blues + depressed + failure + not_worth_living, training)
plot(tree1, type = 'simple')

```


SEXUALITY AND STD AND YOUTH

```{r}

#scale_stop yourself and use birth control once aroused
#plan ahead to have birth control
#resist sexual intercourse without birth control
self_efficacy <- allwave1%>%
  select(AID, H1SE1, H1SE2, H1SE3)

#sex, pregnancy, AIDS/STD test/treatment
#binary_sexual intercourse?
#binary_pregnant? 20.000 missing data
#
Sex_data <- allwave1 %>%
  select(AID, H1CO1,H1FP7,H1HS9)

#first time year, 12000 missing
# binary_birth control, 12 000 missing?,
first_sex <- allwave1%>%
  select(AID, H1CO2Y, H1CO3)

#intercourse without consent
without_consent <- allwave1%>%
  select(AID, H1CO10)

STD <- allwave1%>%
  select(AID, 866:875)%>%
  na_if(7)%>%
  drop_na()%>%
  na_if(6)%>%na_if(8)%>%na_if(9)

Knowledge <- allwave1 %>%
  select(AID, 688:707)%>%
  na_if(6)%>%na_if(7)%>%na_if(8)%>%na_if(9)%>%drop_na()

```

