---
title: "Code 1"
output: html_notebook
---

```{r}
library(plyr)
library(corrplot)
library(gridExtra)
library(ggthemes)
library(caret)
library(MASS)
library(randomForest)
library(party)


```

```{r}

#unload

detach(package : plyr)
detach(package :corrplot)
detach(package :gridExtra)
detach(package :ggthemes)
detach(package :caret)
detach(package :MASS)
detach(package :randomForest)
detach(package :party)
#load

```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(mice)
library(VIM)
```


#Initial dataset 

```{r}
names(final_suicide_data)
main <- final_suicide_data%>%
  select(-AID,-BIO.SEX, -AGE, )

manlog <- glm(suicidal_thoughts~., family = binomial(link = "logit"), data = main)
print(summary(manlog))
a<- anova(manlog, test = "Chisq")
summary(a)
sum(a$Deviance, na.rm = TRUE)
```



#Initial regression for dimension reduction
```{r}
logModelinit <- glm(suicidal_thoughts~ suicidal_surrounding + poor_appetite_feel + shake_blues + depressed + failure+ not_worth_living + suicidal_surrounding + BIO.SEX + alcohol + cigarette + marijuana + pay_attention + feel_close_to_people +  headache + stomach_ache + physically_weak +poor_appetite + trouble_sleeping + moodiness+enjoy_life + inhalant , family = binomial(link = "logit"), data = final_suicide_data)
print(summary(logModelinit))
full_dev <- logModelinit$deviance
12444
anova_init <- anova(logModelinit, test = "Chisq")
anova_init
sum(anova_init$Deviance, na.rm = TRUE)

```

# final_data_to analyze

```{r}
my_data <- final_suicide_data %>% select(BIO.SEX,headache, stomach_ache, physically_weak, poor_appetite, trouble_sleeping, moodiness, pay_attention, feel_close_to_people, suicidal_surrounding, poor_appetite_feel, shake_blues, depressed, failure, enjoy_life, not_worth_living, alcohol, cigarette, marijuana, inhalant, suicidal_thoughts)
```











#Bar plots for categorical variables
```{r}
str(final_suicide_data)

p1<- ggplot(final_suicide_data, aes(x = as.numeric(BIO.SEX))) + ggtitle("Biological Sex") +  scale_x_discrete(name = "", limits = c("Male", "Female"))+
  geom_bar(aes(y = 100*stat(count)/sum(stat(count))), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()


p2<- ggplot(final_suicide_data, aes(x = as.factor(AGE))) + ggtitle("Age") + xlab("Age")+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p3<- ggplot(final_suicide_data, aes(x = as.numeric(headache))) + ggtitle("Headache") + scale_x_discrete(name = "", limits = c("Never", "Just a few times", "about once a week", "almost everyday", "every day"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p4<- ggplot(final_suicide_data, aes(x = as.numeric(stomach_ache))) + ggtitle("Stomachache") + scale_x_discrete(name = "", limits = c("Never", "Just a few times", "about once a week", "almost everyday", "every day"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p5<- ggplot(final_suicide_data, aes(x = as.numeric(physically_weak))) + ggtitle("Physically Weak") + scale_x_discrete(name = "", limits = c("Never", "Just a few times", "about once a week", "almost everyday", "every day"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p6<- ggplot(final_suicide_data, aes(x = as.numeric(tired))) + ggtitle("Tired") + scale_x_discrete(name = "", limits = c("Never", "Just a few times", "about once a week", "almost everyday", "every day"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p7<- ggplot(final_suicide_data, aes(x = as.numeric(wake_up_tired))) + ggtitle("Waking up tired") + scale_x_discrete(name = "", limits = c("Never", "Just a few times", "about once a week", "almost everyday", "every day"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p8<- ggplot(final_suicide_data, aes(x = as.numeric(poor_appetite))) + ggtitle("Poor appetite") + scale_x_discrete(name = "", limits = c("Never", "Just a few times", "about once a week", "almost everyday", "every day"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p9<- ggplot(final_suicide_data, aes(x = as.numeric(trouble_sleeping))) + ggtitle("Trouble sleeping") + scale_x_discrete(name = "", limits = c("Never", "Just a few times", "about once a week", "almost everyday", "every day"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p10<- ggplot(final_suicide_data, aes(x = as.numeric(moodiness))) + ggtitle("Moodiness") + scale_x_discrete(name = "", limits = c("Never", "Just a few times", "about once a week", "almost everyday", "every day"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p11<- ggplot(final_suicide_data, aes(x = as.numeric(risk_seatbelt))) + ggtitle("Wear a seatbelt when riding") + scale_x_discrete(name = "", limits = c("Never", "rarely", "sometimes", "most of the time", "always"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p12<- ggplot(final_suicide_data, aes(x = as.numeric(risk_drunk_drive))) + ggtitle("Drive while drinking") + scale_x_discrete(name = "", limits = c("Never", "rarely", "sometimes", "most of the time", "always"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p13<- ggplot(final_suicide_data, aes(x = as.numeric(miss_school))) + ggtitle("Miss school because of health or emotional problem") + scale_x_discrete(name = "", limits = c("Never", "Just a few times", "about once a week", "almost everyday", "every day"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()


p14<- ggplot(final_suicide_data, aes(x = as.numeric(repeat_grade))) + ggtitle("repeated a grade") + scale_x_discrete(name = "", limits = c("No", "Yes"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p15<- ggplot(final_suicide_data, aes(x = as.numeric(suspension))) + ggtitle("suspended") + scale_x_discrete(name = "", limits = c("No", "Yes"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p16<- ggplot(final_suicide_data, aes(x = as.numeric(expelled))) + ggtitle("expelled") + scale_x_discrete(name = "", limits = c("No", "Yes"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p17<- ggplot(final_suicide_data, aes(x = as.numeric(pay_attention))) + ggtitle("pay attention") + scale_x_discrete(name = "", limits = c("Never", "Just a few times", "about once a week", "almost everyday", "every day"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p18<- ggplot(final_suicide_data, aes(x = as.numeric(homework_done))) + ggtitle("homework done") + scale_x_discrete(name = "", limits = c("Never", "Just a few times", "about once a week", "almost everyday", "every day"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p19<- ggplot(final_suicide_data, aes(x = as.numeric(get_along_with_students))) + ggtitle("get along with students") + scale_x_discrete(name = "", limits = c("Strongly agree", "agree", "neither","disagree", "strongly disagree"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()


p20<- ggplot(final_suicide_data, aes(x = as.numeric(feel_close_to_people))) + ggtitle("feel close to people") + scale_x_discrete(name = "", limits = c("Strongly agree", "agree", "neither","disagree", "strongly disagree"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()


p21<- ggplot(final_suicide_data, aes(x = as.numeric(feel_part_of_your_school))) + ggtitle("feel part of the school") + scale_x_discrete(name = "", limits = c("Strongly agree", "agree", "neither","disagree", "strongly disagree"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p22<- ggplot(final_suicide_data, aes(x = as.numeric(suicidal_thoughts))) + ggtitle("Suicidal thoughts") + scale_x_discrete(name = "", limits = c("No", "Yes"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p23<- ggplot(final_suicide_data, aes(x = as.numeric(suicidal_surrounding))) + ggtitle("Suicidal surrounding") + scale_x_discrete(name = "", limits = c("No", "Yes"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p24<- ggplot(final_suicide_data, aes(x = as.numeric(poor_appetite_feel))) + ggtitle("feel like having poor appetite") + scale_x_discrete(name = "", limits = c("Never or rarely", "sometimes", "a lot of the time", "most of the time or all the time"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p25<- ggplot(final_suicide_data, aes(x = as.numeric(shake_blues))) + ggtitle("feel like you could not shake off the blues") + scale_x_discrete(name = "", limits = c("Never or rarely", "sometimes", "a lot of the time", "most of the time or all the time"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p26<- ggplot(final_suicide_data, aes(x = as.numeric(depressed ))) + ggtitle("felt depressed") + scale_x_discrete(name = "", limits = c("Never or rarely", "sometimes", "a lot of the time", "most of the time or all the time"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p27<- ggplot(final_suicide_data, aes(x = as.numeric(failure ))) + ggtitle("life had been a failure") + scale_x_discrete(name = "", limits = c("Never or rarely", "sometimes", "a lot of the time", "most of the time or all the time"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p28<- ggplot(final_suicide_data, aes(x = as.numeric(Happy ))) + ggtitle("were happy") + scale_x_discrete(name = "", limits = c("Never or rarely", "sometimes", "a lot of the time", "most of the time or all the time"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p29<- ggplot(final_suicide_data, aes(x = as.numeric(lonely ))) + ggtitle("felt lonely") + scale_x_discrete(name = "", limits = c("Never or rarely", "sometimes", "a lot of the time", "most of the time or all the time"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p30<- ggplot(final_suicide_data, aes(x = as.numeric(enjoy_life  ))) + ggtitle("enjoyed life") + scale_x_discrete(name = "", limits = c("Never or rarely", "sometimes", "a lot of the time", "most of the time or all the time"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p31<- ggplot(final_suicide_data, aes(x = as.numeric(sad ))) + ggtitle("felt sad") + scale_x_discrete(name = "", limits = c("Never or rarely", "sometimes", "a lot of the time", "most of the time or all the time"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p32<- ggplot(final_suicide_data, aes(x =as.numeric(not_worth_living))) + ggtitle("felt life was not worth living") + scale_x_discrete(name = "", limits = c("Never or rarely", "sometimes", "a lot of the time", "most of the time or all the time"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()


p33<- ggplot(final_suicide_data, aes(x = as.numeric(alcohol))) + ggtitle("Alcohol") + scale_x_discrete(name = "", limits = c("No", "Yes"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p34<- ggplot(final_suicide_data, aes(x = as.numeric(cigarette))) + ggtitle("Cigarette") + scale_x_discrete(name = "", limits = c("No", "Yes"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p35<- ggplot(final_suicide_data, aes(x = as.numeric(marijuana))) + ggtitle("Marijuana") + scale_x_discrete(name = "", limits = c("No", "Yes"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p36<- ggplot(final_suicide_data, aes(x = as.numeric(cocaine))) + ggtitle("Cocaine") + scale_x_discrete(name = "", limits = c("No", "Yes"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p37<- ggplot(final_suicide_data, aes(x = as.numeric(inhalant))) + ggtitle("Inhalant") + scale_x_discrete(name = "", limits = c("No", "Yes"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p38<- ggplot(final_suicide_data, aes(x = as.numeric(illegal_drug)))+ ggtitle("Illegal drug") + scale_x_discrete(name = "", limits = c("No", "Yes"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p39<- ggplot(final_suicide_data, aes(x = as.numeric(night_away))) + ggtitle("Night away without permission") + scale_x_discrete(name = "", limits = c("No", "Yes"))+
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_linedraw()

map



grid.arrange(p1, p2, p22, p23, ncol = 2)
grid.arrange( p3, p4, p5, p6,p7, ncol =2)
grid.arrange(p8, p9, p10,p13, ncol = 2) 
grid.arrange(p11, p12, p39, ncol = 2)
grid.arrange(p14, p15, p16, p17, p18, ncol = 2)
grid.arrange(p19, p20, p21, ncol = 2)
grid.arrange(p24, p25, p26, p27, p28, p29, ncol = 2)
grid.arrange(p30, p31, p32, ncol = 2)
grid.arrange(p33, p34, p35, p36, p37, p38, ncol = 2)


N = nrow(training)
dim(training);dim(testing)

```



#Start splitting data for k-fold crossvalidation

```{r}
set.seed(78)
intrain <- createDataPartition(my_data$suicidal_thoughts, times = 1, p = .7, list = FALSE)
training <- my_data[intrain,]
testing <- my_data[-intrain,]

```

#logistic regression
```{r}
my_log_model <- glm(suicidal_thoughts ~ .,family = binomial(link = "logit"), data = training)


fitted_results <- predict(my_log_model, newdata = testing, type = 'response')
fitted_results <- as.matrix(fitted_results)
fitted_results <- ifelse(fitted_results>0.5, 1,0)
error <- mean(fitted_results == testing$suicidal_thoughts, na.rm = TRUE)
print(paste('Logistic Regression Accuracy', error))

fitted_table <- as.factor(fitted_results)
confusionMatrix(fitted_table, testing$suicidal_thoughts)

                          
```
```{r}
print("Confusion Matrix for Logistic regression");table(testing$suicidal_thoughts, fitted_results>0.5)

True_negative2 <- 5208
True_Positive2<-  216
False_negative2<-  608
False_positive2<-  117
Actual_true2 <- 824
Actual_false2 <- 5806
predicted_true2 <- 333
Accuracy2 <- round((True_Positive2+True_negative2)/6149, digits = 2)
print(paste('Overall, the predictions are correct', Accuracy2*100,"% of time" ))
error_rate2 <- round((False_negative2 + False_positive2)/6149, digits =2)
print(paste('Overall, the predictions are wrong', error_rate2*100,"% of time" ))
True_positive_rate2 <- round(True_Positive2/Actual_true2, digits = 2)
print(paste('When students actually had suicidal thoughts, it predicts them correctly', True_positive_rate2*100,"% of time" ))
False_positive_rate2 <- round(False_positive2/Actual_false2, digits = 2)
print(paste('When students did not actually have suicidal thoughts, it predicts  they did', False_positive_rate2*100,"% of time" ))
True_negative_rate2<- round(True_negative2/Actual_false2, digits = 2)#sensitivity
print(paste('When students did not actually have suicidal thoughts, it predicts they did not', True_negative_rate2*100,"% of time" ))
precision2 <- round(True_Positive2/predicted_true2,digits =2)#specificity 
print(paste('When the model predicts suicidal thoughts, it is correct', precision2*100,"% of time" ))


```


#random forest and confusion matrix
```{r}
rfModel <- randomForest(suicidal_thoughts ~ . , data = training)


pred_Rf <- predict(rfModel, testing)
confusionMatrix(pred_Rf, testing$suicidal_thoughts)
errorrf <- mean(pred_Rf == testing$suicidal_thoughts, na.rm = TRUE)
print(paste('decision tree Accuracy', errorrf))
plot(rfModel, xaxt = "n")
axis(1, at = seq(0, 500, by = 10),las = 2)
abline(v=60, col = "blue")

plot(rfModel)
lines(x=60, col = "red")
num_tree <- 60
rfModel <- randomForest(suicidal_thoughts ~ . , data = training, ntree = 60, mtry = sqrt(20))

pred_Rf1 <- predict(rfModel, testing)
confusionMatrix(pred_Rf1, testing$suicidal_thoughts)
errorrf1 <- mean(pred_Rf1 == testing$suicidal_thoughts, na.rm = TRUE)
print(paste('random tree Accuracy', errorrf1))

True_negative2 <- 5211
True_Positive2<-  184
False_negative2<-  114
False_positive2<-  117
Actual_true2 <- 824
Actual_false2 <- 5806
predicted_true2 <- 333
Accuracy2 <- round((True_Positive2+True_negative2)/6149, digits = 2)
print(paste('Overall, the predictions are correct', Accuracy2*100,"% of time" ))
error_rate2 <- round((False_negative2 + False_positive2)/6149, digits =2)
print(paste('Overall, the predictions are wrong', error_rate2*100,"% of time" ))
True_positive_rate2 <- round(True_Positive2/Actual_true2, digits = 2)
print(paste('When students actually had suicidal thoughts, it predicts them correctly', True_positive_rate2*100,"% of time" ))
False_positive_rate2 <- round(False_positive2/Actual_false2, digits = 2)
print(paste('When students did not actually have suicidal thoughts, it predicts  they did', False_positive_rate2*100,"% of time" ))
True_negative_rate2<- round(True_negative2/Actual_false2, digits = 2)#sensitivity
print(paste('When students did not actually have suicidal thoughts, it predicts they did not', True_negative_rate2*100,"% of time" ))
precision2 <- round(True_Positive2/predicted_true2,digits =2)#specificity 
print(paste('When the model predicts suicidal thoughts, it is correct', precision2*100,"% of time" ))

```
```{r}

plot(rfModel)
```

```{r}
library(class)
library(dplyr)
library(e1071)
library(FNN)
library(gmodels)
library(psych)
library(caret)
```


#knn and confusion matrix
```{r}

#create feature and target training/test sets
features_train <- my_data[intrain, !(names(my_data) %in% c('suicidal_thoughts'))]
features_test <- my_data[-intrain, !(names(my_data) %in% c('suicidal_thoughts'))]
target_train <- my_data[intrain, "suicidal_thoughts"]
target_test <- my_data[-intrain, "suicidal_thoughts"]

#process each new set
 
preprocess_object <- preProcess(features_train, method = c('scale', 'center', 'knnImpute'))
features_train <- predict(preprocess_object, features_train)
features_test <- predict(preprocess_object, features_test)

#fit the model

knn_fit <- knn3(features_train, target_train, k = 120)
knn_pred <- predict(knn_fit, features_test, type = 'class')

#Calculate error

#error_knn <- mean(ifelse(target_test != knn_pred, 1, 0))

error_knn <- mean(target_test == knn_pred)
print(paste('knn classification Accuracy', error_knn))

confusionMatrix(knn_pred, target_test)

         

```

#decision tree using ctree and confusion matrix

```{r}
dtreeci <- ctree(suicidal_thoughts ~ . , data = training)

pred_tree <- predict(dtreeci, testing)
p1<- predict(dtreeci, training)

tab1 <- table(Predicted = p1, Actual = training$suicidal_thoughts)
tab2 <- table(Predicted = pred_tree, Actual = testing$suicidal_thoughts)

errordt <- mean(pred_tree == testing$suicidal_thoughts, na.rm = TRUE)
#print(paste('decision tree Accuracy', sum(diag(tab2))/sum(tab2)))
print(paste('decision tree Accuracy', errordt))
tab1
plot(dtreeci)

```

#decision tree using rpart and confusion matrix

```{r}
library(rpart)
library(rpart.plot)
trad_dtree <- rpart(suicidal_thoughts~., data = training, method = "class")
p <- predict(trad_dtree, testing, type = "class")
errordt_trad <- mean(p == testing$suicidal_thoughts, na.rm = TRUE)
print(paste('decision tree Accuracy', errordt_trad))

rpart.plot(trad_dtree, type = 4, box.palette = c("red","green"), fallen.leaves = TRUE)

plotcp(trad_dtree)
printcp(trad_dtree)

pruned <- prune(trad_dtree, cp = 0.01)
rpart.plot(pruned, type = 4, box.palette = c("red","green"), fallen.leaves = TRUE)
printcp(pruned)

p1 <- predict(pruned, testing, type = "class")
errordt_pruned <- mean(p1 == testing$suicidal_thoughts, na.rm = TRUE)
print(paste('decision tree Accuracy', errordt_pruned))

confusionMatrix(p1, testing$suicidal_thoughts)

```



#k-fold cross validation for knn
```{r}
accuracy_calc <- function(split_pro, folds, kn) {

  split_index <- createDataPartition(my_data$suicidal_thoughts, times = folds, p = split_pro, list = FALSE) 
  
  accuracy_df <- data.frame(matrix(ncol = 2, nrow = ncol(split_index)))
  colnames(accuracy_df) <- c('test_error', 'fold')
  
  for(i in 1:nrow(accuracy_df)){

    #create feature and target training/test sets
features_train <- my_data[split_index[,i], !(names(my_data) %in% c('suicidal_thoughts'))]
features_test <- my_data[-split_index[,i], !(names(my_data) %in% c('suicidal_thoughts'))]
target_train <- my_data[split_index[,i], "suicidal_thoughts"]
target_test <- my_data[-split_index[,i], "suicidal_thoughts"]

#process each new set

preprocess_object <- preProcess(features_train, method = c('scale', 'center', 'knnImpute'))
features_train <- predict(preprocess_object, features_train)
features_test <- predict(preprocess_object, features_test)

#fit the model

knn_fit <- knn3(features_train, target_train, k = kn)
knn_pred <- predict(knn_fit, features_test, type = 'class')

#Calculate error

accuracy_knn <- mean(target_test == knn_pred)
accuracy_df[i, 'test_error'] <- accuracy_knn
accuracy_df[i,'fold']<- i
  }
  print(mean(accuracy_df$test_error))
  return(accuracy_df)
}

kn_120 <- accuracy_calc(.7, 10, 120)



```

#k-fold cross validation for logistic
```{r}

accuracy_log_calc <- function(split_pro, folds) {

  split_index <- createDataPartition(my_data$suicidal_thoughts, times = folds, p = split_pro, list = FALSE) 
  
  accuracy_df <- data.frame(matrix(ncol = 2, nrow = ncol(split_index)))
  colnames(accuracy_df) <- c('test_error', 'fold')
  
  for(i in 1:nrow(accuracy_df)){

    #create feature and target training/test sets
training <- my_data[split_index[,i],]
testing <- my_data[-split_index[,i],]


#fit the model

my_log_model <- glm(suicidal_thoughts ~ .,family = binomial(link = "logit"), data = training)

fitted_results <- predict(my_log_model, newdata = testing, type = 'response')
fitted_results <- as.matrix(fitted_results)
fitted_results <- ifelse(fitted_results>0.5, 1,0)

#Calculate error
accuracy_log <- mean(fitted_results == testing$suicidal_thoughts, na.rm = TRUE)
accuracy_df[i, 'test_error'] <- accuracy_log
accuracy_df[i,'fold']<- i
  }
  print(mean(accuracy_df$test_error))
  return(accuracy_df)
  
 
}

lg<- accuracy_log_calc(.7, 10)


```
# k fold cross validation rpart decision tree
```{r}

accuracy_dtree_calc <- function(split_pro, folds) {

  split_index <- createDataPartition(my_data$suicidal_thoughts, times = folds, p = split_pro, list = FALSE) 
  
  accuracy_df <- data.frame(matrix(ncol = 2, nrow = ncol(split_index)))
  colnames(accuracy_df) <- c('test_error', 'fold')
  
  for(i in 1:nrow(accuracy_df)){

    #create feature and target training/test sets
training <- my_data[split_index[,i],]
testing <- my_data[-split_index[,i],]


#fit the model

trad_dtree <- rpart(suicidal_thoughts~., data = training, method = "class")
pruned <- prune(trad_dtree, cp = 0.01)


fitted_results <- predict(pruned, testing, type = "class")


#Calculate error
accuracy_dtree <- mean(fitted_results == testing$suicidal_thoughts, na.rm = TRUE)
accuracy_df[i, 'test_error'] <- accuracy_dtree
accuracy_df[i,'fold']<- i
  }
  print(mean(accuracy_df$test_error))
  return(accuracy_df)
 
}
dt <- accuracy_dtree_calc(.7, 10)

```


# k fold cross validation random forest and accuracy rate
```{r}
accuracy_rforest_calc <- function(split_pro, folds) {

  split_index <- createDataPartition(my_data$suicidal_thoughts, times = folds, p = split_pro, list = FALSE) 
  
  accuracy_df <- data.frame(matrix(ncol = 2, nrow = ncol(split_index)))
  colnames(accuracy_df) <- c('test_error', 'fold')
  
  for(i in 1:nrow(accuracy_df)){

    #create feature and target training/test sets
training <- my_data[split_index[,i],]
testing <- my_data[-split_index[,i],]


#fit the model

rfModel <- randomForest(suicidal_thoughts ~ . , data = training, ntree = 60, mtry = sqrt
                        (20), na.action = na.roughfix)


fitted_results <- predict(rfModel, testing)



#Calculate error
accuracy_rforest <- mean(fitted_results == testing$suicidal_thoughts, na.rm = TRUE)
accuracy_df[i, 'test_error'] <- accuracy_rforest
accuracy_df[i,'fold']<- i
  }
  print(mean(accuracy_df$test_error))
  return(accuracy_df)
  
 
}
rf<- accuracy_rforest_calc(.7, 10)
```

#accuracy rate for other algorithm  
```{r}
bknn <- accuracy_calc(.7, 100, 120)


bdt<- accuracy_dtree_calc(.7,100)

blg <- accuracy_log_calc(.7,100)

brf <- accuracy_rforest_calc(.7, 100)

```

```{r}
cknn <- accuracy_calc(.9, 100, 120)


cdt<- accuracy_dtree_calc(.9,100)

clg <- accuracy_log_calc(.9,100)

crf <- accuracy_rforest_calc(.9, 100)
```

```{r}
dknn <- accuracy_calc(.5, 100, 120)

ddt<- accuracy_dtree_calc(.5,100)

dlg <- accuracy_log_calc(.5,100)

drf <- accuracy_rforest_calc(.5, 100)

```

#Visualization
```{r}

ten_fold_graph <- ggplot(kn_120, 
       aes(x = fold, y = test_error, color = "knn")) + 
  geom_line()+
  xlab("Fold") + ylab("Accuracy rate") + ggtitle("Accuracy rate VS 10 Fold accross different models at 70% split")+
  geom_line(aes(x = lg$fold, y = lg$test_error,color = 'logistic regression'))+
  geom_line(aes(x = dt$fold, y = dt$test_error, color = 'decision tree'))+
  geom_line(aes(x = rf$fold, y = rf$test_error, color = 'random forest'))+
    scale_colour_manual("", values = c("logistic regression" = "blue","random forest"="black","decision tree" = "green", "knn" = "red"))
hun_fold_graph<- ggplot(bknn, 
       aes(x = fold, y = test_error, color = "knn")) + 
  geom_line()+
  xlab("Fold") + ylab("Accuracy rate") + ggtitle("Accuracy rate VS 100 Fold accross different models at 70%")+
  geom_line(aes(x = blg$fold, y = blg$test_error,color = 'logistic regression'))+
  geom_line(aes(x = bdt$fold, y = bdt$test_error, color = 'decision tree'))+
  geom_line(aes(x = brf$fold, y = brf$test_error, color = 'random forest'))+
    scale_colour_manual("", values = c("logistic regression" = "blue","random forest"="black","decision tree" = "green", "knn" = "red"))

lar_prop_graph<- ggplot(cknn, 
       aes(x = fold, y = test_error, color = "knn")) + 
  geom_line()+
  xlab("Fold") + ylab("Accuracy rate") + ggtitle("Accuracy rate VS 100 Fold accross different models at 90% split")+
  geom_line(aes(x = clg$fold, y = clg$test_error,color = 'logistic regression'))+
  geom_line(aes(x = cdt$fold, y = cdt$test_error, color = 'decision tree'))+
  geom_line(aes(x = crf$fold, y = crf$test_error, color = 'random forest'))+
    scale_colour_manual("", values = c("logistic regression" = "blue","random forest"="black","decision tree" = "green", "knn" = "red"))

mid_prop_graph<- ggplot(cknn, 
       aes(x = fold, y = test_error, color = "knn")) + 
  geom_line()+
  xlab("Fold") + ylab("Accuracy rate") + ggtitle("Accuracy rate VS 100 Fold accross different models at 50% split")+
  geom_line(aes(x = dlg$fold, y = dlg$test_error,color = 'logistic regression'))+
  geom_line(aes(x = ddt$fold, y = ddt$test_error, color = 'decision tree'))+
  geom_line(aes(x = drf$fold, y = drf$test_error, color = 'random forest'))+
    scale_colour_manual("", values = c("logistic regression" = "blue","random forest"="black","decision tree" = "green", "knn" = "red"))

mid_prop_graph
lar_prop_graph
hun_fold_graph 
ten_fold_graph


```


#ndi data #not used 

```{r}
death_cause <- ndi%>%
  select(AID, NDICOD07)%>%
  na_if(8)%>%na_if(9)%>%drop_na()

death_cause<-death_cause%>%
  mutate(suicide = as.factor(ifelse(death_cause$NDICOD07==2,1,0 )))%>%
  select(-NDICOD07)
         

table(death_cause$suicide)

suicide_data <- left_join(death_cause, final_suicide_data, by = "AID")
suicide_data <- suicide_data%>%
  select( suicide,BIO.SEX,headache, stomach_ache, physically_weak, poor_appetite, trouble_sleeping, moodiness, pay_attention, feel_close_to_people, suicidal_surrounding, poor_appetite_feel, shake_blues, depressed, failure, enjoy_life, not_worth_living, alcohol, cigarette, marijuana, inhalant, suicidal_thoughts)


set.seed(78)
intrain1 <- createDataPartition(suicide_data$suicide, times = 1, p = .7, list = FALSE)
training1 <-suicide_data[intrain1,]
testing1 <- suicide_data[-intrain1,]


dim(training1);dim(testing1)
my_log_model1 <- glm(suicide ~ suicidal_surrounding + suicidal_thoughts,family = binomial(link = "logit"), data = training1)



print(summary(my_log_model1))
an5 <- anova(my_log_model1, test = "Chisq")
sum(an4$Deviance, na.rm = TRUE)

fitted_results2 <- predict(my_log_model1, newdata = testing1, type = 'response')
fitted_results <- as.matrix(fitted_results)
fitted_results <- ifelse(fitted_results>0.5, 1,0)
error <- mean(fitted_results == testing$suicidal_thoughts, na.rm = TRUE)
print(paste('Logistic Regression Accuracy', error))

```